---
title: Livability scores of select neighborhoods world-wide
subtitle: Comparing US cities and some similar cities based on their Livability score
author: Fadeli Mohamed Elwardi
date: 
fontsize: 10pt
output:
 beamer_presentation:
  theme: "Luebeck"
  colortheme: "orchid"
  fonttheme: "professionalfonts"
# keep_tex: true
# toc: true
  slide_level: 3
  #includes:
   #in_header: /home/elwardi/Course/presentations/header.tex
 #after_body: ~/Dropbox/teaching/table-of-contents.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
use_python("usr/local/bin/python3")
```

### Liveablity-based clustering is important

- You know, figuring out which cities/neighborhoods are easier or harder to live in!

\medskip

- It's also crucial for predicting whether a city will be leaving its current cluster of 
  similar cities in the near features
  - Obviously, one should leave if the current city will get harder to live in
  - Where to head?
  - Well, to the easy-living city whose probability of leaving its cluster is minimal!

\medskip

- But this project takes care only of the first step: The clustering of cities.

### Data acquisition and cleaning

- A modified version of the scheme shown at **[This webpage](https://www.areavibes.com/methodology/)**
  is used to estimate Livability Aspects.

\medskip

- For this project, we rely on two main sources of data:

1. **[A huge data set of NYC neighborhood stats](https://www1.nyc.gov/assets/doh/downloads/excel/nta/nta-metadata.xlsx)**
2. **[Foursquare API](https://develop.foursquare.com)**

- To get arround 390 rows of data (Including some other cities worldwide that were added manually)

\medskip

- Data cleanup is performed while getting the Foursquare data
  - No latitude/longitude data for each neighborhood were supplied
  - Instead, search "near" these location in Foursquare and retrieve the coordinates

### Data collected from NYC neighborhoods Database

```{python, echo=FALSE}
import pandas as pd
data = pd.read_csv('/tmp/full_data.csv')
cols = ['Poverty','Violent Crime','Property Crime','EduLessThanHS',\
'Crowded Housing','Health Ins', 'Unemployment Rate']
```
```{r, echo=FALSE}
library('ggplot2')
par(mar=c(10,8.5,1.5,1.5))
boxplot(py$data[py$cols],
        main = 'Data from NYC neighborhoods data set',
        xlab = "Index Value",
        col = "orange",
        border = "darkblue",
        horizontal = TRUE,
        las=1,
        notch = TRUE)
```

### Data collected from Foursquare

```{python, echo=FALSE}
cols = ['Groceries', 'Food & Drink', 'Shopping', 'Schools', 'Entertainment',
       'Fitness Facilities', 'Transportation', 'Libraries',
       'Goods and Services']
```
```{r, echo=FALSE}
library('ggplot2')
par(mar=c(10,8.5,1.5,1.5))
boxplot(py$data[py$cols],
        main = 'Data from Foursquare',
        xlab = "Index Value",
        col = "orange",
        border = "darkblue",
        horizontal = TRUE,
        las=1,
        notch = FALSE)
```

### How these data features correlate with each other?

```{r, echo=FALSE}
df = py$data
names(df) = make.names(names(df))
pairs(py$data[,2:10], pch = 19, lower.panel = NULL, bty="n", 
      col="darkblue", cex.labels= 0.8, cex=.4, cex.lab=.2, 
      cex.axis=.5, las=2,
      main='Correlation between data features collected from Foursquare')
```

### How these data features correlate with each other?

```{r, echo=FALSE}
df = py$data
names(df) = make.names(names(df))
pairs(py$data[,11:16], pch = 19, lower.panel = NULL, bty="n", col="darkblue", 
      cex.labels= 1.0, cex=.4, cex.lab=.2, cex.axis=.5,las=2,
      main='Correlation between features collected from NYC neighborhood stats')
```

### What to estimate from the data?

Each Data feature is multiplied by a contribution coefficient for each Livability Aspect:

\scriptsize 

```{python, echo=FALSE}
import pandas as pd
scoring = {
    'Amenities':{
        'Groceries': 0.17,
        'Food & Drink': 0.17,
        'Shopping': 0.26,
        'Schools': 0.12,
        'Entertainment': 0.16,
        'Fitness Facilities': 0.06,
        'Transportation': 0.03,
        'Libraries': 0.03
    },
    'Cost of Living':{
        'Goods and Services': 0.3,
        'Groceries': 0.1,
        'Transportation':0.1,
        'Crowded Housing': 0.25,
        'Health Ins': 0.15,
        'Poverty': 0.1
    },
    'Crime':{
        'Violent Crime': 0.65,
        'Property Crime': 0.35
    },
    'Education':{
        'EduLessThanHS': 0.85,
        'Schools': 0.25
    },
    'Employment':{
        'Unemployment Rate': 0.1, 
    }
}
sc = pd.DataFrame.from_dict(scoring)
sc.fillna(value=0.0, inplace=True)
```

```{r, echo=FALSE}
knitr::kable(py$sc)
```

\normalsize

### How the estimated indices correlate?

```{python, echo=FALSE}
for i in range(len(sc.transpose())):
    tmp = pd.DataFrame(sc.transpose().iloc[[i]].values*data[data.columns[0:16]], columns=data.columns[0:16])
    data[sc.transpose().index[i]] = pd.DataFrame.sum(tmp,axis=1)
```
```{r, echo=FALSE}
df = py$data
names(df) = make.names(names(df))
pairs(py$data[,21:25], pch = 19, lower.panel = NULL, bty="n", col="darkblue", 
      cex.labels= 1.2, cex=.4, cex.lab=.2, cex.axis=.5,las=2,
      main='Correlation between main features of Livability Index')
```

### Classification - NYC neighborhoods

- NYC neighborhoods look very similar Amenities-wise

\begin{figure}[b]  \centering 
  \includegraphics[width=0.8\textwidth]{/home/elwardi/coursera/Capstone/Coursera_Capstone/01-cluster-amenities.png}
	\caption{NYC neighborhoods clustering based on Amenities presence}\label{im01}
\end{figure}


### Classification - NYC neighborhoods

- They can also be clustered into 2 groups (All 5 Livability Aspects):

\begin{figure}[b]  \centering 
  \includegraphics[width=0.75\textwidth]{/home/elwardi/coursera/Capstone/Coursera_Capstone/01-cluster-all-nyc.png}
	\caption{NYC neighborhoods clustering based on Livability Aspects}\label{im02}
\end{figure}

### Classification - Broader view

- The cities around New York are more diverse in terms of Livability similarity:

\begin{figure}[b]  \centering 
  \includegraphics[width=0.8\textwidth]{/home/elwardi/coursera/Capstone/Coursera_Capstone/01-zoomout-nyc.png}
	\caption{Clustering of some US cities based on Livability Aspects}\label{im03}
\end{figure}

### Conclusion

- Analyzed the livability index of several important locations
- Identified which cities and neighborhoods look most similar when compared by the five main aspects of the Livability score

\medskip

- Acquiring Premium Foursquare account can greatly improve the model!

\medskip

- Can now predict when a city may leave its current cluster!
  - By repeatedly applying the model at different (past) time frames